### 2023-05-01 draft1.Rmd
updated

---
title: "FM project"
output: html_document
date: "2023-04-25"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## Analysing U.S. energy exports

### Introduction

Energy commodities play a vital role in the global economy as they are an essential ingredient for economic growth and development. Energy, among others, is required for transportation, manufacturing, heating, and cooling, and many other critical activities that drive economic activity.

The production, consumption, and trade of energy commodities also have significant impacts on the global economy. The prices of energy commodities are influenced by supply and demand factors, as well as geopolitical tensions, weather events, and technological advances.

Lately, society has faced two significant challenges: firstly, the COVID-19 pandemic, and secondly, the war in Ukraine. These have created unprecedented changes in the energy market, and understanding the impact of these events is crucial for policymakers, analysts, and investors.

The objective of this study is to examine the time-series data on energy commodity exports from the United States in the energy market. The United States has been chosen because it is one of the leading producers and exporters of energy commodities, and its export trends have significant implications for the global energy market. Also, the United States serves as a significant alternative for EU countries to turn to after the EU's imposition of sanctions on Russia.

Moreover, the main goals of this study are to:

* Identify the main patterns and trends in the selected series, including any seasonality, cyclicity, or trends, and determine if any series are white noise.

* Use time series decomposition methods to identify the effects of different components, such as trend, seasonality, and residual, on the series and determine if any adjustments or transformations are needed.

* Apply different forecasting methods, including simple forecasting methods, exponential smoothing, and ARIMA models, to forecast future values of the selected series and evaluate the accuracy of these methods.

* Explore other forecasting methods not covered in class, such as dynamic regression models, Prophet model, and Neural Network Autoregression, to compare their performance with the methods covered in class.

The data sets utilized in this study consist of monthly observations of Country’s reliance on Russian oil products, USA’s natural gas exports by country, USA’s total crude oil exports by destination, and USA’s petroleum products exports by destination.

Four fallowing tsibbles were compiled out of four data sets:

* Country’s reliance on Russian oil products (Monthly)

## Data preparation and overview

Every data set was

Our primary source of data is the U.S. Energy Information Administration (EIA), an independent organization that collects, analyzes, and disseminates data and information related to energy production, consumption, and distribution in the United States. Also, for the purpose of this study we are using data from the International Energy Agency (IEA), which  is an intergovernmental organization that was established in 1974 to promote global cooperation on energy issues.

In case of data on reliance on Russian Oil the dataset contains monthly total oil imports from Russia as a percentage of total oil imports for a selected group of European countries belonging to the Organization for Economic Co-operation and Development (OECD) from January 1990 to October 2022. The dataset includes the date, country name, and the percentage of total oil imports that came from Russia for each observation.

```{r, Reliance on Russian oil_tsibble example, echo=FALSE, results='markup',warning=FALSE,message=FALSE}

library(tidyverse)
library(readr)
library(lubridate)
library(tsibble)

oecd_europe <- c("Austria", "Belgium", "Czech Republic", "Denmark", "Estonia",
                 "Finland", "France", "Germany", "Greece", "Hungary", "Iceland",
                 "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg",
                 "Netherlands", "Norway", "Poland", "Portugal", "Slovak Republic",
                 "Slovenia", "Spain", "Sweden", "Switzerland", "Turkey",
                 "United Kingdom")

raw_rel_rusoil <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Monthly_Reliance_on_Russian_Oil.csv", skip = 1)

rel_rusoil <- raw_rel_rusoil %>% 
  pivot_longer(cols = -...1,
               names_to = 'Date',
               values_to = 'Total Oil Imports from Russia / Total Oil Imports (%)')

rel_rusoil <- rel_rusoil %>% 
  rename(Country = ...1) %>% 
  mutate(Date = lubridate::dmy(Date)) %>% 
  select("Date","Country","Total Oil Imports from Russia / Total Oil Imports (%)")

ts_rel_rusoil <- rel_rusoil %>% 
  as_tsibble(
    index = Date,
    key = Country)

oecd_relrus <- ts_rel_rusoil %>%
  filter(Country %in% oecd_europe)

oecd_relrus %>% slice_sample(n=10)

```

In case of other three data sets

```{r,Crude oil petroleum products and nat gas tsibble examples ,echo=FALSE, results='markup',warning=FALSE,message=FALSE}
library(tidyverse)
library(readr)
library(lubridate)
library(tsibble)

################################################################################

#Total Crude Oil Exports by Destination

raw_crude_oil <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Total%20Crude%20Oil%20Exports%20by%20Destination_EIA.csv", skip = 2,)

#Leaving only country names in column names

names(raw_crude_oil)[-2] <- str_replace_all(names(raw_crude_oil)[-2],
                                            c("U.S. Exports to " = "",
                                              " of Crude Oil \\(Thousand Barrels\\)" = ""))

#Pivoting (switching) columns starting from 2 column to rows

crude_oil <- raw_crude_oil %>% 
  pivot_longer(cols = -Date,
               names_to = 'Destination country',
               values_to = 'Amount of Crude Oil (Thousand Barrels)')

#Filtering out values where Date is 'NA'

crude_oil <- crude_oil %>%
  drop_na(Date) %>%
  mutate(Date = as.Date(Date))

#Fixing date format, replacing 'na' values with 0

crude_oil$Date <- as.Date(crude_oil$Date)
crude_oil$`Amount of Crude Oil (Thousand Barrels)` <- replace_na(crude_oil$`Amount of Crude Oil (Thousand Barrels)`, 0)

#Setting up the tsibble

ts_crude_oil <- crude_oil %>%
  mutate(Date = yearmonth(Date)) %>% 
  as_tsibble(
    index = Date, 
    key = `Destination country`
  )

#OECD Europe Countries vector

oecd_europe <- c("Austria", "Belgium", "Czech Republic", "Denmark", "Estonia",
                 "Finland", "France", "Germany", "Greece", "Hungary", "Iceland",
                 "Ireland", "Italy", "Latvia", "Lithuania", "Luxembourg",
                 "Netherlands", "Norway", "Poland", "Portugal", "Slovak Republic",
                 "Slovenia", "Spain", "Sweden", "Switzerland", "Turkey",
                 "United Kingdom")

#filtering out only relevant values

oecd_crude_oil <- ts_crude_oil %>%
  filter(`Destination country` %in% oecd_europe) %>% 
  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))

#Total Oil Products Exports by Destination

raw_oil_prod <- read_csv("https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Total%20Oil%20Products%20Exports%20by%20Destination_EIA.csv",  skip = 2,)

names(raw_oil_prod)[-2] <- str_replace_all(names(raw_oil_prod)[-2],
                                           c("U.S. Exports to " = "",
                                             " of Total Petroleum Products \\(Thousand Barrels\\)" = ""))
oil_prod <- raw_oil_prod %>% 
  pivot_longer(cols = -Date,
               names_to = 'Destination country',
               values_to = 'Amount of total Petroleum Porducts (Thousand Barrels)')

oil_prod$`Amount of total Petroleum Porducts (Thousand Barrels)` <- replace_na(oil_prod$`Amount of total Petroleum Porducts (Thousand Barrels)`, 0)

oil_prod <- oil_prod %>% 
  drop_na(Date) %>% 
  mutate(Date = as.Date(Date))

ts_oil_prod <- oil_prod %>%
  mutate(Date = yearmonth(Date)) %>% 
  as_tsibble(
    index = Date, 
    key = `Destination country`
  )

oecd_oil_prod <- ts_oil_prod %>% 
  filter(`Destination country` %in% oecd_europe) %>% 
  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))

oecd_oil_prod %>% slice_sample(n=10)

################################################################################
#Total Natural Gas Exports by Destination

raw_nat_gas <- read.csv('https://raw.githubusercontent.com/edb-313/Energy-trade-forecasting/main/Data/csv/Natural%20Gas%20Exports%20by%20Country_EIA.csv',header=TRUE , check.names=FALSE, skip = 2)

# drop columns 2-3, 6-8, 54 and 57-72
raw_nat_gas <- subset(raw_nat_gas, select = -c(2:3, 6:8,54, 57:72))
raw_nat_gas[is.na(raw_nat_gas)] <- 0

#cimbining exports to the same countries

nat_gas <- raw_nat_gas %>%
  mutate(Mexico = `U.S. Natural Gas Pipeline Exports to Mexico (MMcf)` + `Liquefied U.S. Natural Gas Exports by Truck to Mexico (Million Cubic Feet)`,
         Canada = `U.S. Natural Gas Pipeline Exports to Canada (MMcf)` + `Liquefied U.S. Natural Gas Exports by Truck to Canada (Million Cubic Feet)`)

#removing duplicate columns

nat_gas <- nat_gas[, !(names(nat_gas) %in% c('U.S. Natural Gas Pipeline Exports to Mexico (MMcf)',
                                             'Liquefied U.S. Natural Gas Exports by Truck to Mexico (Million Cubic Feet)',
                                             'U.S. Natural Gas Pipeline Exports to Canada (MMcf)',
                                             'Liquefied U.S. Natural Gas Exports by Truck to Canada (Million Cubic Feet)'))]

#cleaning column names

colnames(nat_gas) <- gsub(pattern = ".*(to\\s)", "", colnames(nat_gas))
colnames(nat_gas) <- gsub("\\(.*?\\)", "", colnames(nat_gas))

#pivoting data
gastemp <- nat_gas %>% 
  pivot_longer(cols = -Date,
               names_to = 'Destination country',
               values_to = 'Amount of Natural gas (MMcf)')

#removing n/as in date column, fixing date formatting, fixing values (romoving space in the end)
library(stringr) #for triming the strings

gastemp <- gastemp %>% 
  drop_na(Date) %>% 
  mutate(Date = as.Date(Date)) %>% 
  mutate(`Destination country` = str_trim(`Destination country`))

#summing duplicates

gastemp <- gastemp %>%
  group_by(Date, `Destination country`) %>%
  summarize_at(vars(`Amount of Natural gas (MMcf)`), sum) %>%
  ungroup() %>% 
  drop_na(Date)

#creating a tsibble

ts_nat_gas <- gastemp %>%
  mutate(Date = yearmonth(Date)) %>% 
  as_tsibble(
    index = Date,
    key = `Destination country`
  )

#filtering only relevant data

oecd_nat_gas <- ts_nat_gas %>%
  filter(`Destination country` %in% oecd_europe) %>% 
  filter(Date >= as.Date("2013-01-15"), Date <= as.Date("2023-01-15"))

oecd_nat_gas %>% slice_sample(n=10)
```

**Placeholder for methodologies**

## 2. Time series graphics

Done

## 3. Time series decomposition

WIP

## 4. The forecaster’s toolbox

Almost done, needs bit more input

## 5. Exponential smoothing

Yet to be started

## 6. ARIMA

Yet to be started

## 7. Other forecasting methods

Yet to be started